
\documentclass [11pt, proquest, natbib] {uwthesis}[2020/02/24]
\usepackage{CJKutf8}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{listings}
\newcommand\todo[1]{\textcolor{red}{TODO: #1}}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsthm}
\mathchardef\mhyphen="2D

% \newcommand{\egg}{\texorpdfstring{\MakeLowercase{\texttt{egg}}}{\texttt{egg}}\xspace}
% \newcommand{\Egg}{\texorpdfstring{\MakeLowercase{\texttt{egg}}}{\texttt{egg}}\xspace}
\newcommand{\egg}{\texorpdfstring{\MakeLowercase{\textsc{egg}}}{\texttt{egg}}\xspace}
\newcommand{\Egg}{\texorpdfstring{\MakeLowercase{\textsc{egg}}}{\texttt{egg}}\xspace}
\newcommand{\egraphs}{\mbox{e-graphs}\xspace}
\newcommand{\egraph}{\mbox{e-graph}\xspace}
\newcommand{\Egraph}{\mbox{E-graph}\xspace}
\newcommand{\Egraphs}{\mbox{E-graphs}\xspace}
\newcommand{\eclass}{\mbox{e-class}\xspace}
\newcommand{\Eclass}{\mbox{E-class}\xspace}
\newcommand{\enode}{\mbox{e-node}\xspace}
\newcommand{\eclasses}{\mbox{e-classes}\xspace}
\newcommand{\enodes}{\mbox{e-nodes}\xspace}
\newcommand{\Enodes}{\mbox{E-nodes}\xspace}
\newcommand{\Ematching}{\mbox{E-matching}\xspace}
\newcommand{\ematching}{\mbox{e-matching}\xspace}
\newcommand{\child}{\textit{child}}
\newcommand{\sym}{\textit{symbol}}
\newcommand{\eclassid}{\textit{eclass-id}}
\newcommand{\id}{\textit{Id}}



%
% The following line would print the thesis in a postscript font 

% \usepackage{natbib}
% \def\bibpreamble{\protect\addcontentsline{toc}{chapter}{Bibliography}}

\setcounter{tocdepth}{1}  % Print the chapter and sections to the toc
 

% ==========   Local defs and mods
%

% --- sample stuff only -----
% These format the sample code in this document

\usepackage{alltt}  % 
\newenvironment{demo}
  {\begin{alltt}\leftskip3em
     \def\\{\ttfamily\char`\\}%
     \def\{{\ttfamily\char`\{}%
     \def\}{\ttfamily\char`\}}}
  {\end{alltt}}
 
% metafont font.  If logo not available, use the second form
%
% \font\mffont=logosl10 scaled\magstep1
\let\mffont=\sf
% --- end-of-sample-stuff ---
 



\begin{document}
 
% ==========   Preliminary pages
%
% ( revised 2012 for electronic submission )
%

\prelimpages
 
%
% ----- copyright and title pages
%
\Title{Faster and Worst-Case Optimal E-Matching}
\Author{Yihong Zhang}
\Year{June, 2021}
\Degree{Bachelor of Science \\with Departmental Honors}
\School{University of Washington}
\Program{Computer Science \& Engineering,}
\Programtext{}

\Chair{Name of Chairperson}{Title of Chair}{Department of Chair}
\Signature{First committee member}
\Signature{Next committee member}
\Signature{etc}

\copyrightpage

\titlepage  


\setcounter{page}{-1}
\abstract{%
An \egraph is a data structure that efficiently represents sets of congruent terms \citep{eqsatthesis}. 
It has a long history of applications in automated theorem proving \citep{eqsatthesis, efficient-ematching, cc-in-tt,proof-producing,simplify} and is re-purposed in the last decade for program optimization, known as equality saturation \citep{eqsat, denaili, spores, semsearch,herbie, egg}. 
A fundamental query operation on e-graph is e-matching \citep{efficient-ematching, simplify}, which finds the set of terms in the e-graph matching a given pattern. 
E-matching is the bottleneck of many equality saturation--based program optimizers \citep{spores,herbie,egg} and is a central procedure in state-of-the-art SMT solvers, including \texttt{Z3} \citep{Z3} and \texttt{CVC4} \citep{CVC4}.
Therefore, the performance of e-matching is critical.
Several optimizations are proposed for e-matching \citep{efficient-ematching,simplify}.
However, to our knowledge, existing e-matching algorithms are based on na\"ive backtracking, which is suboptimal in many cases. 
In particular, they do not exploit the equality constraints implied by multi-occurrences of a variable, which is nonetheless common in practice.

To tackle this inefficiency, we propose to take a relational view of e-graphs.
Under this view, e-matching corresponds naturally to a restricted class of relational queries, known as conjunctive query. 
By treating e-matching as an instance of conjunctive query, we benefit from decades of researches from the database community. 
In particular, by using the recently discovered generic join algorithm, our e-matching algorithm guarantees worst-case optimality and achieves more than 400$\times$ speed-up in our preliminary experiments.

This thesis is based on an extended abstract to appear at PLDI 2021 Student Research Competition and a forthcoming paper.
}
 
%
% ----- contents & etc.
%
\tableofcontents
\listoffigures
%\listoftables  % I have no tables
 
\include{chapters/0-acknowledgments}

%
% end of the preliminary pages
 
 
 
%
% ==========      Text pages
%

\textpages

\include{chapters/1-introduction}
\include{chapters/2-background}
\include{chapters/3-relational-ematch}
\include{chapters/5-discussions}
\include{chapters/4-optimizations}
\include{chapters/6-evaluations}
\include{chapters/7-conclusions}
\include{chapters/8-futureworks}
 
% ========== Chapter 2
 
\printendnotes

%
% ==========   Bibliography
%
\medskip

\bibliographystyle{ACM-Reference-Format}
\bibliography{uwthesis}


\end{document}
