\acknowledgments{% \vskip2pc
  % {\narrower\noindent
  Along my journey of academics and life, I have owed thanks to many people. They teach me, help me, encourage me, and care about me. I thank them for everything they did for me.
  
  I would like to first thank my thesis advisor Professor Zachary Tatlock, who is always encouraging and gives me great help during my thesis work. I also would like to thank Professor Dan Suciu, who was kind enough for recognizing my worth and introduced me to my first research on genetic algorithms for counterfactual explanations. I thank Max Schleich for his kind guidance on the counterfactual explanation project and my academic journey. I also would like to thank James Wilcox. I worked both as a student of and a teaching assistant for Zach and James, and I deeply appreciate their dedication to students and enthusiasm to creating a wonderful learning environment for all.
  
  I would like to thank my collaborators and mentors Remy Wang and Max Willsey on the work described in this thesis. It is always my pleasure to discussing ideas with them and learning from their insights, which I deeply appreciate.  It is them who make this project possible.
  
  I would like to thank my roommate Mike He, Jiuru Li, and Bruce Li for their care and tolerance throughout my collegiate life. We share interests on many things. In particular, Mike and I have the same enthusiasm in programming languages and I learned so many things from him.
  
  I want to thank my parents, who brought me to the world and supports me, emotionally and financially, to study abroad at the States. I recognize it is such a privilege for me to study here at UW, and this is impossible without my parents. 
  
  I want to thank Jiawei Huang, my love, who is always here when I am down and celebrates for my achievements. It's my biggest fortune to meet her and fall in love with her.
  
  Finally, I want to thank my maternal grandfather Peihua Zhao and grandmother Jinhua Yu. Words cannot express how deeply owed I am to them. They taught me to be a righteous human being and were always proud of what I did. My heart broke when they passed away in 2016 and in 2020, but they keep appearing in my dreams. This thesis is dedicated to them.
  % \par}
}

\dedication{\begin{center}To my maternal grandmother and grandfather,

I miss you. 

\begin{CJK*}{UTF8}{gbsn}
外公外婆，我想念你们。
\end{CJK*}
\end{center}
}
